Ensemble Method for Time Series Forecasting

### 2018-01-13 paper.tex
Correcting lexical and grammar errors

%Single insitute
\author{Aleksandar Bachvarov \and Atanas Dimitrov}
%If there are too many authors, use \authorrunning
%\authorrunning{First Author et al.}
\institute{Karlsruhe Institute of Technology}

@@

\begin{abstract}
Abstract goes here
\end{abstract}

@@

\section{Introduction}\label{sec:intro}

@@

	\subsection{Definition}
		\subsubsection{Time series}
		 \hspace{1cm}\\\\A time series is sequence of data points, which are ordered and indexed by time. The data presented in the series could be any variable capable for observation over time. Some examples of time series include the price of a specific stock, the height level of a river and the national birth rate over time. Time series could be presented by multiple ways: as a list of values, as a graph, as a bar chart etc. On  \hyperref[fig:bitcoin]{Fig. \ref{fig:bitcoin}} you can see a graph of the bitcoin price time series over the last year. Beside investing, time series is used in many other fields like e.g. statistics, insurance, weather forecasting, astronomy, applied  science and many more.

@@

		\subsubsection{Forecasting} \hspace{1cm}\\\\Time series forecasting represents the process of predicting the future values of the observed variable based on the already available data. In order to achieve that, dependencies among the available time series must be found, which then approximate the possible future values. Various methods were used over the years. They are in general divided  in two groups: linear and non-linear methods. The most significant of them will be wider presented  further in this paper. It's important to be noted that time series forecasting is just forecasting - the purpose is that the forecast is as close as possible to the future value, but there is no warranty for that. Generally applies that the deviation of the forecast increases with his time distance.\\\\ Time series could be not-stationary (volatile) or stationary. In other words - to have higher or lower degree of stationarity. The distribution of the stationary one depends much more on a long term trend while the distribution of the not-stationary one - on occasional events. Example for an extremely stationary time series is the coin flipping over time, example for an extremely volatile one - the bitcoin price. Because of his characteristics the volatile time series is much harder to predict. His wide spread and importance led to the development of more complex (non-linear) methods for doing that. Those methods include and are based on some machine learning techniques like e.g. bagging and boosting and also on key machine learning tools like artificial neural networks. Time series forecasting is one of the classic and at the same time very important areas of the machine learning.

@@

In the recent years, as many useful applications of machine learning have been developed and have become part of our everyday life, his significance increased. We hear more and more often the term on the news or read about it on the Internet. In many cases machine learning is used as a synonym for artificial intelligence (AI), which is incorrect. AI is field of computer science, which is studying and  applying approaches for making computers being able to execute cognitive tasks\cite{AIvsML}. Machine learning is one of those approaches and it's by far the most successful (widely used, with the most applications) one of them\cite{quinlan1986induction}. His subject is the studying and the computer modelling of learning processes.\cite{Michalski1983}. \\\\ One common way to create a new machine learning process is to research the human learning mechanisms, to adapt them and then accordingly to simulate them  with algorithms. The artificial neural networks (commonly utilized machine learning tool) are example for that and they will be discussed thoroughly further in the paper. It is  restrictive to believe that the learning patterns that come from the nature are the only possible way for acquiring knowledge. Reserchers also try to manifacture their own ones. By doing this main criteria are the methods'  generality and performance rather than their psychological explanation\cite{Michalski1983}.\\\\ 

@@

\end{enumerate}\hspace{1cm}\\Different amount of inference means that the learning system analyses the data more or less and makes more or less interdependencies between it. Example for no inference is the simple remembering of facts with no thoughts about them and any other already existing knowledge. Example for middle range inference is the learning from teacher or book that provides the learner positive and negative examples. The learner analyses the input data and makes some conclusion about it based mainly on his already existing knowledge. Example for extremely high level of inference is the learning from observation and the discovery. There is no source of certain information to learn from and the learner should discover it alone\cite{Michalski1983}. The high and the low level of inference of learning methods has both their pluses and minuses - it's much harder and slower to learn something just by yourself, but therefore you reach deeper understanding of the topic. When the learner is a real person and not a computer algorithm, it could be even dangerous for him. My driving instructor used to tell me that I can always learn to drive by myself, but to achieve it, I must break at least three cars. In order to prevent that he was there to teach me.\\\\

@@

We know from the near past and the present  (and for sure will hear more about it in the future) that the machine learning has a lot of applications. Some popular application domains of the learners are the self driving cars, speech and face recognition, the expert systems and of course - the subject of our paper - time series forecasting.  \\\\As next on this paper follow some presentation of key tools and techniques in machine learning at all and in the field of time series forecasting specifically.

\paragraph{Decision trees}
 A decision tree could be defined as form of classification or regression rule that a learner has inducted after his feeding with data. Let's assume we have a set of objects. All these objects have attributes, which  have values. Besides that we also know that each object belongs to one specific class and there are at least two different classes. From that data a decision tree could be inducted. Latter we can use that decision tree for classification - defining to which of the classes any object belongs. Under \enquote{any} here are all objects from the same type as the input objects to be understand. The regression decision tree is similar to the classification one - the difference is that the delight output isn't class label as by classification, but it's a continuous value. For better understanding we will stay on the classification in the further explanations\cite{quinlan1986induction}.\\\\It must be said that the decision tree is used for predictions. In that matter there is no 100\% warranty that the prediction true. In order to increase the probability for correct forecast, we must first select as less redundant data as possible. With other words that selected objects must cover the whole spectrum of different objects, but with no duplicates. When this condition isn't fulfilled, we have the so called underfitting\cite{quinlan1986induction}.\\\\
Even without underfitting there are still chance that the decision tree doesn't return the correct class label. This may be caused by overfitting - when the decision tree corresponds too much to the training data, but isn't abstract enough (doesn't return correct class labels for other data). In order to fix that problem many ensemble techniques could be used. One of them is the so called \enquote{Random forest}. A random forest is a set of different decision trees that we have generated through e.g. sampling. In order to determine the output value, we chose the mode of the selected classes from all trees in the forest\cite{breiman2001random}.\\\\
An example decision tree is given on \hyperref[fig:decisionTree]{Fig. \ref{fig:decisionTree}}. It represents the classification of whether or not you should give credit to someone. The nodes are shown as conditions for the values of the object's attributes. The edges are signed with the possible values of the attributes and route the path of the selection. The leafs are the different class label. On the example we see that you will probably get credit, if you are first over 40 years old and own a house, because it will be mortgaged or if you earn enough money to cover your debt.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{decisionTree}
\caption{Decision tree about whether to give credit to someone or not}
\label{fig:decisionTree}
\end{figure}

@@

The artificial neural networks are one of the widely employed tools in the machine learning. They simulate the learning and predicting activity of the brain. Similar to the process of teaching humans the artificial neural nets are trained with pairs of data - input variable and his corresponding result value according to the task. Then the net should be able to guess the output value of random input value that wasn't part of the training set. \\\\The  fundamental part of an artificial neural network is the model of the neuron. It consists of three elements\cite{haykin2009neural}:
\vspace{-\topsep}
\begin{enumerate}
\item set of synapses. Each synapse has a weight $w_{kj}$.  When passing through the synapse, the input signal $x_j$ is multiplied by his weight. The weight can hold positive and negative values.
\item an adder. Sums the multiplied input signals. 
\item an activation function. The sum of the adder's value and the bias $b_k$ goes as his argument. The function returns the output signal of the neuron. His purpose is to restrict the output signal between specific values e.g. -1 and 1.
\end{enumerate}
\vspace{-\topsep}

@@

\begin{equation}
y_k = φ(v_k)
\end{equation}Shortly explained the multiplication of all input signals $x_j$ with the weights $w_{kj}$ of their synapses are getting summed. This sum is then biased with $b_k$ and the result is put into the activation function $φ(.)$, which produce the output signal $y_k$.\\\\
The are some possibilities for activation function, but the mostly used are the sigmoid functions. A sigmoid function has a \enquote{S} shaped graph and and return value from 0 to 1 or from -1 to 1. One example is the logistic function.
\begin{equation}
φ(v) = \frac{1}{1 + exp(e^{-x})}
\end{equation}
One particular reason why the sigmoid functions are mostly employed as activation functions is the fact that they are differentiable. This quality is latter important for the backpropagation. \\\\The simplest neural net consist from only one neuron. Further, we can have many neurons in one single layer. In that  case the net is called single-layer feedforward network. If the net consists from more than one layer, then it's called multilayer feedforward network. Multilayer networks have an input layer, an output layer and could also a hidden layer or many of them. The ones with many hidden layers are called deep networks\cite{haykin2009neural}.

@@

\hspace{1cm}\\ The ultimate goal of training each neural network is the adjusting of all weight values to such way that by passing an unknown input signal to the network, it will return a correct output signal. In order to achieve that a technique called backpropagation is used. Simply explained we need to pass an input signal through the network. Then we calculate the difference between the real expected value and the output value. With a special loss function and the difference as argument we calculate the new weight for each synapse and then adjust it by propagating the network backwards\cite{rumelhart1986learning}.\\\\ In order to employ the backpropagation the model have to fulfil two conditions. First, the activation function must be differentiable. The logistic function fulfils that. Second, there has to be no addition of $u_k$ with the bias, but at the same time the effect of the bias must be kept. That's achieved by using an extra dummy synapse with weight $w_{k0} = b_k$ and an input signal equals 1 or -1\cite{haykin2009neural}.
		 \paragraph{Ensemble Techniques}
In machine learning and statistics, combining forecasts from different models are recognised as one of the most successful approaches to prediction tasks and are used to obtain better predictive performance, to enhance accuracies and produce overall improved results. 
%(e.g. Dietterich (2000); Brown and Kuncheva (2010))
\\\\However, forecast combination is a challenging task. 
The idea is to find a suitable hypothesis that will make adequate predictions for a particular problem and then ensemble methods are essentially strategies for creating multiple hypotheses and combining them to generate a better hypothesis. This usually produces more accurate solutions than a single hypothesis would. Using adequate ensemble techniques guarantees aggregation of the forecasting strengths, diminution of their weaknesses and efficient error reduction.\\\\There is a recent suggestion that there is an ideal number of component classifiers called "the law of diminishing returns in ensemble construction". 
%(https://www.researchgate.net/profile/Hamed_R_Bonab)
Because the number of component classifiers has a high impact on the accuracy of the prediction, having more or less than this ideal number would deteriorate the accuracy. \\\\We distinguish one ensemble approach from another by the way they manipulate the training data, by the way of selection of individual hypotheses and by the way in how those are aggregated to the final decision. 
%Chapter 13 ensemble methonds
In this particular section we will cover some commonly used ensemble approaches, namely Bagging, AdaBoost and Stacking(?).

\subsubsection{1. Bagging}
%chapter 13 cite all refferences.
\hspace{1cm}\\\\This method is introduced by Breiman [ref] and is used for training each learner with help from a bootstrap sample from a training set S and then combining each output by uniform averaging. To generate each individual predictor, no information about the predictions of the other models is integrated. For this particular reason this technique is viewed as an algorithm with only an implicit search for diversity. By averaging a set of predictors which are not trained by the same bootstrap sample, there might be a reduction of the influence of the outliers [ref]. Taking that into consideration, resampling seems to have a similar effect to that of robust M-estimators in statistics [ref],  meaning the influence of sample points is restricted using suitable loss functions. Therefore, it is possible that this approach will significantly reduce the variance of the bagged prediction, especially for unstable base learners [ref]. \\\\Because of the fact that traditional bootstrap breaks the sequence of data points, there is a possibility for a bad generalization in time series problems. Inoue and Kilian adjust to the situation by adapting the way that Bagging construct its training samples [ref]. In order to simultaneously capture the reliance between the examples and to encourage diversity,  the authors have to use resampling blocks of instances from the original training set with replacement. This process is called "block bootstrap", and the block size is chosen to capture the dependence in error term [ref].
\\\\Bagging, according to a simulation study,  significantly reduces the out-of-sample mean squared error of the predictions. By giving a weight for each example that resemble the number of times the point is selected in the bootstrap sample, the training samples in [ref] maintain the temporal arrangement of the points. In this manner and example is weighted to zero if it has not been selected in the sample. Conclusively, the output of the ensemble for a single instance is computed as a weighted average, where the weight of each predictor is the weighted root mean squared error (WRMSE). [ref]

\subsubsection{2. AdaBoost}
%chapter 13 cite all ref
\hspace{1cm}\\\\This method is introduced by Freud and Schapire in 1997 [ref]. The idea is to train a single predictor to minimize the residual of the previous ensemble in each round, which can be described as a sequentially coupled approach fitting an additive model through a forward stagewise strategy [ref]. The theoretical background of the algorithm is based on the "PAC" learning model [59] and its purpose is to create a method that combines a set of weak learners to generate a strong learning algorithm. This approach offers good generalization ability but on the other hand, because of the fact that the learners are coupled sequentially it is computationally expensive. \\\\Maintaining a sampling distribution over the training set is the main purpose behind AdaBoost. With the start of the algorithm, all weights are set to $w^1_m = 1/N, m = 1,..., M$. Here,  $w^n_m$ is the weight of the $m$-th example at round $n$.  The weak learner is forced to focus on the misclassified examples of the training set at each round of the algorithm, because of the specific way in which the weights of the incorrectly classified examples are increased. Hence, instead of focusing on the correctly classified examples, the algorithm concentrates on the "hard" examples, making it possible for AdaBoost to outperform Bagging in normal scenarios. Nevertheless, under noisy data the low generalization capacity of AdaBoost makes the robustness of Bagging generally superior. \\\\ In the field of time series, Shresta and Solomatine introduce AdaBoost.RT [61]. Here the absolute relative error (ARE) for the learner $n$ is computed by considering the error over a threshold $φ$. The weights of the instances with errors lower than $φ$ are reduced by the updating rule, and afterwards normalized in order to make them a distribution. Conclusively, F is obtained as the weighted average of the individual learners, with weights equal to - $log(β_n)$. \\\\For tide levels forecasting, Canestrelli uses Drucker's AdaBoost.R2(62).  This approach quantifies the loss for each training sample with the help of one of the three possible functions: 
\begin{enumerate}
\item Absolute percentage error (APE):
\begin{equation}
φ(v) = \frac{|f_n(x_m)-y_m|}{max_{m=1,...,M} |f_n(x_m)-y_m|}
\end{equation}
\item Quadratic:
\begin{equation}
w^n_m = (f_n(x_m)-y_m)^2 max_{m=1,...,M} (f_n(x_m)-y_m)^2
\end{equation}
\item Exponential:
\begin{equation}
1 - exp\frac{-|f_n(x_m)-y_m|}{max_{m=1,...,M} |f_n(x_m)-y_m|}
\end{equation}
\end{enumerate}

\hspace{1cm}\\\\According to the weight distribution, the confidence of the predictor is computed as a function of the weighted average error. In order to increase their chances to appear in the following training sample, the instances with higher errors increase their weights. Then the output of the ensemble is computed as the weighted median. \\\\ For recurrent neural networks as the base learner, Assaad et al. adapts this method [63]. The following changes are present: first, the example errors are weighted for updating the RNN weights instead of resampling. The motivation behind this to keep the temporal dependence of the training data. Second, the weight of the wrong modeled examples is further increased using an additional parameter. \\\\ As an adaptation of AdaBoost.R2 technique with Genetic Programming (GP) as the base learner, Souza et al. [64] proposes the BCC algorithm. This essentially computes the error of each example with the help of the exponential loss function.  And by including a multiplicative factor which depends on the correlation among the learner predictions and the true targets, the algorithm updates the weights. \\\\ Finally, as a combination of Ellman Newtorks using a weighted linear function, Goh et al. [65] introduces the Modified AdaBoost. On top of that the error of each instance is computed by taking into account two loss functions. These are widely used for analysing drug dissolution profiles. [chapter 13] \\\\ The goal of AdaBoost, according to the gradient descent perspective of AdaBoost [8,66], is to add the i-th learner in the ensemble which minimizes
\begin{equation}
(β_n, f_n)=argmin_{(β,f)} \sum_{m=1}^{M} l(y_m, F(x_m)).
\end{equation}
Here $F(x_m)=F_{n-1}(x_m)+ β_n f_n (x_m)$. With each following step, the algorithm selects the hypothesis $f_n$ which minimizes Eq. (8) over the sample distribution $w^n$. Then it performs a line search along the direction. From this perspective, Buhlmann and Yu [67] introduce a gradient boosting method with the quadratic loss function which is adapted for time seies forecasting for load forecasting [68], finance [69], economics [70] and production [71].

\subsubsection{3. Stacking}
\hspace{1cm}\\\\ Stacking, also known as stacked generalization, first introduced by David H. Wolpert [?] in 1992, is a strategy for minimizing the generalization error rate of one or more generalizers. The method works by deducting the biases of the generalizers with respect to a provided learning set. For instance it uses the guess of the original generalizers as input when taught with part of the learning set and trying to guess the rest of it, and the correct guess for output. The difference between using multiple generalizers and using a single generalizer is that in the first case, stacked generalization can be considered as a more sophisticated version of cross-validation for it exploits a better strategy than cross-validation's crude winner-takes-all [?] for combining the individual generalizers. In the second case, when used with a single generalizer, stacked generalization is a schema for estimating the error of a generalizer, trained on a particular learning set and then asked a particular question. \\\\Stacking has a distinguishing feature and that is that the information fed up the net of generalizers comes from multiple partitionings of the original set, feeding information from one set of generalizers to another, then making the final guess. Each of these generalizers splits up that learning set into two subsets. Collecting information about the biases of the generalizing behaviour of the original generalizers with respect to the learning set is the purpose of these pairs of subsets. This collected bias information is the one fed up the net. One can conclude that stacked generalization as a means of estimating and correcting for the biases of the constituent generalizers with respect to the corresponding learning set, can be expected to reduce the generalization error rate for many generalization problems, according to the experiments done in [?].

>>>>

%Single insitute
\author{Aleksandar Bachvarov \and Atanas Dimitrov}
%If there are too many authors, use \authorrunning
%\authorrunning{First Author et al.}
\institute{
Karlsruhe Institute of Technology\\
\email{aleksandar.bachvarov@gmail.com}, \email{atanas\_sz@abv.bg}
}

@@

\begin{abstract}
Time series forecasting is an important field of study in the machine learning. It's influential in the modern insecure world, for it gives approximated data of future events (e.g. vulcanic eruption, birth and death rate estimation, demand for energy), which leads to proper preparations for them. In this paper an innovative ensemble method for time series forecasting is introduced, as well as the time series forecasting evolution throughout history and its machine learning background. The method is based on an ensemble strategy that combines arbitrated by a meta-learning model forecasters. The model is dynamically adapting over time by being constantly fed with new data, which improves the accuracy of the predictions and the overall performance.
\end{abstract}

@@

\section{Introduction}\label{sec:intro}
This paper is the essential part of a seminar work related to Mobile Computing with particular topic \enquote{Ensemble Method for Time Series Forecasting}. It's an assignment in our B.Sc. Computer Science studies at Karlsruhe Institute of Technology. We chose the topic for a number of reasons: One of our personal interests are finance and stock trading. Hence, as the stock prices (and prices at all) are time series, we wanted to dive into the topic of time series predicting tasks and doing so to educate ourselves about financial assets price forecasting. Secondly, the topic is narrowly connected with machine learning - a field of computer science with great significance and endless potential, in which we are motivated to develop ourselves. Through our researches we will get to know more or less its basics.

@@

	\subsection{Definition}
		\subsubsection{Time series}
		 \hspace{1cm}\\\\A time series is sequence of data points, which are ordered and indexed by time. The data presented in the series could be any variable capable for observation over time. Some examples of time series include the price of a specific stock, the height level of a river and the national birth rate over time. Time series could be presented by multiple ways: as a list of values, as a graph, as a bar chart etc. On  \hyperref[fig:bitcoin]{Fig. \ref{fig:bitcoin}} you can see a graph of the bitcoin price time series over the last year. Besides in investing, time series is used in many other fields like e.g. statistics, insurance, weather forecasting, astronomy, applied  science etc.

@@

		\subsubsection{Forecasting} \hspace{1cm}\\\\Time series forecasting represents the process of predicting the future values of the observed variable based on the already available data. In order to achieve that, dependencies among the available time series must be found, which then approximate the possible future values. Various methods were used over the years. They are in general divided  in two groups: linear and non-linear methods. The most significant of them will be wider presented  further in this paper. It's important to be noted that time series forecasting is just forecasting - the purpose is that the forecast is as close as possible to the future value, but there is no warranty for that. Generally the deviation of the forecast increases with his time distance.\\\\ Time series could be non-stationary (volatile) or stationary. In other words - to have higher or lower degree of stationarity. It appears that the distribution of the stationary one depends much more on a long term trend while the distribution of the not-stationary one - on occasional events. Example for an extremely stationary time series is the coin flipping over time, example for an extremely volatile one - the bitcoin price. Because of its characteristics the volatile time series is much harder to predict. Its wide spread and importance led to the development of more complex (non-linear) methods for doing that. Those methods include and are based on some machine learning techniques like e.g. bagging and boosting and also on key machine learning tools like artificial neural networks. Time series forecasting is one of the classic and at the same time very important areas of the machine learning.

@@

In the recent years, as many useful applications of machine learning have been developed and have become part of our everyday life, its significance increased. We hear more and more often the term on the news or read about it on the Internet. In many cases machine learning is used as a synonym for artificial intelligence (AI), which is inexact. AI is field of computer science, which is studying and  applying approaches for making computers being able to execute cognitive tasks\cite{AIvsML}. Machine learning is one of those approaches and it's by far the most successful (widely used, with the most applications) one of them\cite{quinlan1986induction}. Its subject is the studying and the computer modelling of learning processes.\cite{Michalski1983}. \\\\ One common way to create a new machine learning process is to research the human learning mechanisms, to adapt them and then accordingly to simulate them  with algorithms. The artificial neural networks (commonly utilized machine learning tool) are example for that and they will be discussed thoroughly further in the paper. It is  restrictive to believe that the learning patterns that come from the nature are the only possible way for acquiring knowledge. Reserchers also try to manifacture their own ones. By doing this main criteria are the methods'  generality and performance rather than their psychological explanation\cite{Michalski1983}.\\\\ 

@@

\end{enumerate}\hspace{1cm}\\Different amount of inference means that the learning system analyses the data more or less and makes more or less interdependencies between it. Example for no inference is the simple remembering of facts with no thoughts about them and any other already existing knowledge. Example for middle range inference is the learning from teacher or book that provides the learner positive and negative examples. The learner analyses the input data and makes some conclusion about it based mainly on its already existing knowledge. Example for extremely high level of inference is the learning from observation and the discovery. There is no source of certain information to learn from and the learner should discover it alone\cite{Michalski1983}. The high and the low level of inference of learning methods has both their pluses and minuses - it's much harder and slower to learn something single-handed, but because of that you reach deeper understanding of the topic.\\\\

@@

We know from the near past and the present  (and for sure will hear more about it in the future) that the machine learning has a lot of applications. Some popular application domains are the self driving cars, speech and face recognition, the expert systems and of course - the subject of our paper - time series forecasting.  \\\\As next on this paper follow some presentation of key tools and techniques in machine learning at all and in the field of time series forecasting specifically.

\paragraph{Decision trees}
 A decision tree could be defined as form of classification or regression rule that a learner has inducted after its feeding with data. Let's assume we have a set of objects. All these objects have attributes, each of which  has values. Besides that, we also know that each object belongs to one specific class and there are at least two different classes. From that data a decision tree could be inducted. Later we can use that decision tree for classification - defining to which of the classes any object belongs. Under \enquote{any} here are all objects from the same type as the input objects to be understand. The regression decision tree is similar to the classification one - the difference is that the delight output isn't class label as by classification, but it's a continuous value. For better understanding we will stay on the classification in the further explanations\cite{quinlan1986induction}.\\\\It must be said that the decision tree is used for predictions. In that matter there is no 100\% warranty that the prediction is true. In order to increase the probability for correct forecast, we must first select as less redundant data as possible. With other words that selected objects must cover the whole spectrum of different objects, but with no duplicates. When this condition isn't fulfilled, we have the so called underfitting\cite{quinlan1986induction}.\\\\
Even without underfitting there are still chance that the decision tree doesn't return the correct class label. This may be caused by overfitting - when the decision tree corresponds too much to the training data, but isn't abstract enough (doesn't return correct class labels for other data). In order to fix that problem many ensemble techniques could be used. One of them is the so called \enquote{Random forest}. A random forest is a set of different decision trees that we have generated through e.g. sampling. In order to determine the output value, we chose the mode of the selected classes from all trees in the forest\cite{breiman2001random}.\\\\
An example decision tree is given on \hyperref[fig:decisionTree]{Fig. \ref{fig:decisionTree}}. It represents the classification of whether or not one should give credit to someone. The nodes are shown as conditions for the values of the object's attributes. The edges are signed with the possible values of the attributes and route the path of the selection. The leafs are the different class label. On the example we see that one will probably get credit, if he is over 40 years old and owns a house, because it will be mortgaged or if he earns enough money to cover his debt.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{decisionTree}
\caption{Decision tree about choosing whether to give credit to someone or not}
\label{fig:decisionTree}
\end{figure}

@@

The artificial neural networks are one of the widely employed tools in the machine learning. They simulate the learning and predicting activity of the brain. Similar to the process of teaching humans, the artificial neural nets are trained with pairs of data - input variable and their corresponding result value according to the task. Then the net should be able to guess the output value of random input value that wasn't part of the training set. \\\\The  fundamental part of an artificial neural network is the model of the neuron. It consists of three elements\cite{haykin2009neural}:
\vspace{-\topsep}
\begin{enumerate}
\item set of synapses. Each synapse has a weight $w_{kj}$.  When passing through the synapse, the input signal $x_j$ is multiplied by its weight. The weight can hold positive and negative values.
\item an adder. Sums the multiplied input signals. 
\item an activation function. The sum of the adder's value and the bias $b_k$ goes as its argument. The function returns the output signal of the neuron. Its purpose is to restrict the output signal between specific values e.g. -1 and 1.
\end{enumerate}
\vspace{-\topsep}

@@

\begin{equation}
y_k = φ(v_k)
\end{equation}Shortly explained the multiplication of all input signals $x_j$ with the weights $w_{kj}$ of their synapses are getting summed. This sum is then biased with $b_k$ and the result is put into the activation function $φ(.)$, which produce the output signal $y_k$.\\\\
The are some possibilities for activation function, but the mostly used are the sigmoid functions. A sigmoid function has a \enquote{S} shaped graph and and returns value from 0 to 1 or from -1 to 1. One example is the logistic function.
\begin{equation}
φ(v) = \frac{1}{1 + exp(e^{-x})}
\end{equation}
One particular reason why the sigmoid functions are mostly employed as activation functions is the fact that they are differentiable. This quality is later important for the backpropagation. \\\\The simplest neural net consists from only one neuron. Further, we can have many neurons in one single layer. In that  case the net is called single-layer feedforward network. If the net consists from more than one layer, then it's called multilayer feedforward network. Multilayer networks have an input layer, an output layer and could also have a hidden layer or many of them. The ones with many hidden layers are called deep networks\cite{haykin2009neural}.

@@

\hspace{1cm}\\ The ultimate goal of training each neural network is the adjusting of all weight values in such a way that by passing an unknown input signal to the network, it will return a correct output signal. In order to achieve that, a technique called backpropagation is used. Simply explained, we need to pass an input signal through the network. Then we calculate the difference between the real expected value and the output value. With a special loss function and the difference as argument we calculate the new weight for each synapse and then adjust it by propagating the network backwards\cite{rumelhart1986learning}.\\\\ In order to employ the backpropagation the model have to fulfil two conditions. First, the activation function must be differentiable. The logistic function fulfils that. Second, there has to be no addition of $u_k$ with the bias, but at the same time the effect of the bias must be kept. That's achieved by using an extra dummy synapse with weight $w_{k0} = b_k$ and an input signal equals 1 or -1\cite{haykin2009neural}.
		 \paragraph{Ensemble Techniques}
In machine learning and statistics, combining forecasts from different models are recognised as one of the most successful approaches to prediction tasks and are used to obtain better predictive performance, to enhance accuracies and produce overall improved results. 
%(e.g. Dietterich (2000); Brown and Kuncheva (2010))
\\\\However, forecast combination is a challenging task. 
The idea is to find a suitable hypothesis that will make adequate predictions for a particular problem. Ensemble methods are essentially strategies for creating multiple hypotheses and combining them to generate a better hypothesis. This usually produces more accurate solutions than a single hypothesis would. Using adequate ensemble techniques guarantees aggregation of the forecasting strengths, diminution of their weaknesses and efficient error reduction.\\\\There is a recent suggestion that there is an ideal number of component classifiers called \enquote{the law of diminishing returns in ensemble construction}. 
%(https://www.researchgate.net/profile/Hamed_R_Bonab)
Because the number of component classifiers has a high impact on the accuracy of the prediction, having more or less than this ideal number would deteriorate the accuracy. \\\\We distinguish one ensemble approach from another by the way they manipulate the training data, by the way of selection of individual hypotheses and by the way in how those are aggregated to the final decision. 
%Chapter 13 ensemble methonds
In this particular section we will cover some commonly used ensemble approaches, namely Bagging, AdaBoost and Stacking(ref).

\subsubsection{1. Bagging}
%chapter 13 cite all refferences.
\hspace{1cm}\\\\This method is introduced by Breiman [ref] and is used for training each learner with the help of a bootstrap sample from a training set S and then combining each output by uniform averaging. To generate each individual predictor, no information about the predictions of the other models is integrated. For this particular reason this technique is viewed as an algorithm with only an implicit search for diversity. By averaging a set of predictors which are not trained by the same bootstrap sample, there might be a reduction of the influence of the outliers [ref]. Taking that into consideration, resampling seems to have a similar effect to that of robust M-estimators in statistics [ref],  meaning the influence of sample points is restricted using suitable loss functions. Therefore, it is possible that this approach will significantly reduce the variance of the bagged prediction, especially for unstable base learners [ref]. \\\\Because of the fact that traditional bootstrap breaks the sequence of data points, there is a possibility for a bad generalization in time series problems. Inoue and Kilian adjust to the situation by adapting the way that Bagging construct its training samples [ref]. In order to simultaneously capture the reliance between the examples and to encourage diversity,  the authors have to use resampling blocks of instances from the original training set with replacement. This process is called \enquote{block bootstrap}, and the block size is chosen to capture the dependence in error term [ref].
\\\\Bagging, according to a simulation study,  significantly reduces the out-of-sample mean squared error of the predictions. By giving a weight for each example that resemble the number of times the point is selected in the bootstrap sample, the training samples in ...[ref] maintain the temporal arrangement of the points. In this manner, an example is weighted to zero if it has not been selected in the sample. Conclusively, the output of the ensemble for a single instance is computed as a weighted average, where the weight of each predictor is the weighted root mean squared error (WRMSE). [ref]

\subsubsection{2. AdaBoost}
%chapter 13 cite all ref
\hspace{1cm}\\\\This method is introduced by Freud and Schapire in 1997 [ref]. The idea is to train a single predictor to minimize the residual of the previous ensemble in each round, which can be described as a sequentially coupled approach fitting an additive model through a forward stagewise strategy [ref]. The theoretical background of the algorithm is based on the \enquote{PAC} learning model [59] and its purpose is to create a method that combines a set of weak learners to generate a strong learning algorithm. This approach offers good generalization ability, but on the other hand, because of the fact that the learners are coupled sequentially it is computationally expensive. \\\\Maintaining a sampling distribution over the training set is the main purpose behind AdaBoost. With the start of the algorithm, all weights are set to $w^1_m = 1/N, m = 1,..., M$. Here,  $w^n_m$ is the weight of the $m$-th example at round $n$.  The weak learner is forced to focus on the misclassified examples of the training set at each round of the algorithm, because of the specific way in which the weights of the incorrectly classified examples are increased. Hence, instead of focusing on the correctly classified examples, the algorithm concentrates on the \enquote{hard} examples, making it possible for AdaBoost to outperform Bagging in normal scenarios. Nevertheless, under noisy data the low generalization capacity of AdaBoost makes the robustness of Bagging generally superior. \\\\ In the field of time series, Shresta and Solomatine introduce AdaBoost.RT [61]. Here the absolute relative error (ARE) for the learner $n$ is computed by considering the error over a threshold $φ$. The weights of the instances with errors lower than $φ$ are reduced by the updating rule, and afterwards normalized in order to make them a distribution. Conclusively, F is obtained as the weighted average of the individual learners, with weights equal to - $log(β_n)$. \\\\For tide levels forecasting, Canestrelli uses Drucker's AdaBoost.R2(62).  This approach quantifies the loss for each training sample with the help of one of the three possible functions: 
\begin{enumerate}
\item Absolute percentage error (APE):
\begin{equation}
φ(v) = \frac{|f_n(x_m)-y_m|}{max_{m=1,...,M} |f_n(x_m)-y_m|}
\end{equation}
\item Quadratic:
\begin{equation}
w^n_m = (f_n(x_m)-y_m)^2 max_{m=1,...,M} (f_n(x_m)-y_m)^2
\end{equation}
\item Exponential:
\begin{equation}
1 - exp\frac{-|f_n(x_m)-y_m|}{max_{m=1,...,M} |f_n(x_m)-y_m|}
\end{equation}
\end{enumerate}

\hspace{1cm}\\\\According to the weight distribution, the confidence of the predictor is computed as a function of the weighted average error. In order to increase their chances to appear in the following training sample, the instances with higher errors increase their weights. Then the output of the ensemble is computed as the weighted median. \\\\ For recurrent neural networks as the base learner, Assaad et al. adapts this method [63]. The following changes are present: first, the example errors are weighted for updating the RNN weights instead of resampling. The motivation behind this is to keep the temporal dependence of the training data. Second, the weight of the wrong modeled examples is further increased using an additional parameter. \\\\ As an adaptation of AdaBoost.R2 technique with Genetic Programming (GP) as the base learner, Souza et al. [64] proposes the BCC algorithm. This essentially computes the error of each example with the help of the exponential loss function.  And by including a multiplicative factor which depends on the correlation among the learner predictions and the true targets, the algorithm updates the weights. \\\\ Finally, as a combination of Ellman Newtorks using a weighted linear function, Goh et al. [65] introduces the Modified AdaBoost. On top of that, the error of each instance is computed by taking into account two loss functions. These are widely used for analysing drug dissolution profiles. [chapter 13] \\\\ The goal of AdaBoost, according to the gradient descent perspective of AdaBoost [8,66], is to add the i-th learner in the ensemble which minimizes
\begin{equation}
(β_n, f_n)=argmin_{(β,f)} \sum_{m=1}^{M} l(y_m, F(x_m)).
\end{equation}
Here $F(x_m)=F_{n-1}(x_m)+ β_n f_n (x_m)$. With each following step, the algorithm selects the hypothesis $f_n$ which minimizes Eq. (8) over the sample distribution $w^n$. Then it performs a line search along the direction. From this perspective, Buhlmann and Yu [67] introduce a gradient boosting method with the quadratic loss function which is adapted for time seies forecasting for load forecasting [68], finance [69], economics [70] and production [71].

\subsubsection{3. Stacking}
\hspace{1cm}\\\\ Stacking, also known as stacked generalization, first introduced by David H. Wolpert [?] in 1992, is a strategy for minimizing the generalization error rate of one or more generalizers. The method works by deducting the biases of the generalizers with respect to a provided learning set. For instance, it uses the guess of the original generalizers as input (when taught with part of the learning set and trying to guess the rest of it) and the correct guess for output. The difference between using multiple generalizers and using a single generalizer is that in the first case, stacked generalization can be considered as a more sophisticated version of cross-validation for it exploits a better strategy than cross-validation's crude winner-takes-all [?] for combining the individual generalizers. In the second case, when used with a single generalizer, stacked generalization is a schema for estimating the error of a generalizer, trained on a particular learning set and then asked a particular question. \\\\Stacking has a distinguishing feature and that is that the information fed up the net of generalizers comes from multiple partitionings of the original set, feeding information from one set of generalizers to another, then making the final guess. Each of these generalizers splits up that learning set into two subsets. Collecting information about the biases of the generalizing behaviour of the original generalizers with respect to the learning set is the purpose of these pairs of subsets. This collected bias information is the one fed up the net. One can conclude that stacked generalization, as a means of estimating and correcting for the biases of the constituent generalizers with respect to the corresponding learning set, can be expected to reduce the generalization error rate for many generalization problems, according to the experiments done in [?].
